---
title: "Risk Stratification of Practices using Clustering"
format:
  html:
    code-fold: true
    warning: false
    toc: true
    toc-location: left
    number-sections: true
---

```{r}
#| label: library-imports
#| echo: false

library(tidyverse)
library(ggplot2)
library(readxl)


```

```{r}
#| label: setup
#| echo: false


source("composite_scores_calcs.R")

data <- read_xlsx("../data/data.xlsx")

gp_domain_composite_score <- calc_domain_composite_scores(
  data = data |> filter(indicator_level == "GP Level"),
  indicator_id_col = "indicator_id",
  polarity_col = "indicator_polarity",
  domain_col = "domain",
  geography_code_col = "geography_code",
  value_col = "indicator_value",
  exclude_indicators = 7,
  minimum_indicators = 1L,
  minimum_completeness = 0.75
)

gp_cluster_data <- gp_domain_composite_score |>
  select(
    geography_code,
    domain,
    domain_standardised_score
  ) |>
  pivot_wider(
    names_from = domain,
    values_from = domain_standardised_score
  )

```

# Introduction

# Data Exploration

## Distribution of standardised composite domain scores

```{r}
#| label: hist-distribution-domain-scores
#| fig_cap: "Distribution of standardised domain composite scores"
#| fig-height: 10
#| fig-width: 12

domain_order <- c(
  "Acute Coronory Syndrome",
  "Atrial Fibrillation",
  "Stroke",
  "Cholesterol",
  "Hypertension",
  "Heart Failure",
  "Diabetes"
)

# Create facet labels
domain_labels <- gp_domain_composite_score |>
  distinct(
    domain,
    indicators_expected
  ) |>
  mutate(
    domain = factor(
      domain,
      levels = domain_order
    ),
    domain_label = paste0(
      domain,
      "\n(",
      indicators_expected,
      if_else(
        indicators_expected == 1,
        " indicator)",
        " indicators)"
      )
    )
  ) |>
  arrange(domain) |>
  mutate(
    domain_label = factor(
      domain_label,
      levels = domain_label
    )
  )

gp_domain_composite_score |>
  filter(!is.na(domain_standardised_score)) |>
  left_join(
    domain_labels,
    by = c("domain", "indicators_expected")
  ) |>
  ggplot(
    aes(x = domain_standardised_score)
  ) +
  geom_histogram(bins = 20) +
  facet_wrap(
    ~ domain_label,
    scales = "free_y"
  ) +
  labs(
    title = "Distribution of standardised domain scores",
    subtitle = "Number of indicators contributing to each domain shown in brackets",
    x = "Standardised domain score",
    y = "Number of practices"
  ) +
  theme_minimal()

# domain_order <- c(
#   "Acute Coronory Syndrome",
#   "Atrial Fibrillation",
#   "Stroke",
#   "Cholesterol",
#   "Hypertension",
#   "Heart Failure",
#   "Diabetes"
# )
# 
# # Create facet labels
# domain_labels <- gp_domain_composite_score |>
#   distinct(
#     domain,
#     indicators_expected
#   ) |>
#   mutate(
#     domain = factor(
#       domain,
#       levels = domain_order
#     ),
#     domain_label = paste0(
#       domain,
#       "\n(",
#       indicators_expected,
#       if_else(
#         indicators_expected == 1,
#         " indicator)",
#         " indicators)"
#       )
#     )
#   )
# 
# plot_data <- gp_domain_composite_score |>
#   filter(
#     !is.na(domain_standardised_score)
#   ) |>
#   left_join(
#     domain_labels,
#     by = c(
#       "domain",
#       "indicators_expected"
#     )
#   )
# 
# library(patchwork)
# library(grid)
# 
# p1 <- plot_data |>
#   filter(
#     domain %in% c(
#       "Acute Coronory Syndrome",
#       "Atrial Fibrillation",
#       "Stroke"
#     )
#   ) |>
#   ggplot(
#     aes(x = domain_standardised_score)
#   ) +
#   geom_histogram(bins = 20) +
#   facet_wrap(
#     ~ domain_label,
#     nrow = 1,
#     scales = "free_y"
#   ) +
#   labs(
#     x = NULL,
#     y = NULL
#   ) +
#   theme_minimal()
# 
# 
# p2 <- plot_data |>
#   filter(
#     domain %in% c(
#       "Cholesterol",
#       "Hypertension"
#     )
#   ) |>
#   ggplot(
#     aes(x = domain_standardised_score)
#   ) +
#   geom_histogram(bins = 20) +
#   facet_wrap(
#     ~ domain_label,
#     nrow = 1,
#     scales = "free_y"
#   ) +
#   labs(
#     x = NULL,
#     y = NULL
#   ) +
#   theme_minimal()
# 
# 
# p3 <- plot_data |>
#   filter(
#     domain %in% c(
#       "Diabetes",
#       "Heart Failure"
#     )
#   ) |>
#   ggplot(
#     aes(x = domain_standardised_score)
#   ) +
#   geom_histogram(bins = 20) +
#   facet_wrap(
#     ~ domain_label,
#     nrow = 1,
#     scales = "free_y"
#   ) +
#   labs(
#     x = NULL,
#     y = NULL
#   ) +
#   theme_minimal()
# 
# combined_plot <- (p1 / p2 / p3) +
#   plot_annotation(
#     title = "Distribution of standardised domain scores",
#     subtitle = "Number of indicators contributing to each domain shown in brackets"
#   )
# 
# gridExtra::grid.arrange(
#   combined_plot,
#   left = grid::textGrob(
#     "Number of practices",
#     rot = 90,
#     gp = grid::gpar(fontsize = 12)
#   ),
#   bottom = grid::textGrob(
#     "Standardised domain score",
#     gp = grid::gpar(fontsize = 12)
#   )
# )
# 
# ggsave(
#   "plots/domain_distributions.png",
#   width = 12,
#   height = 10,
#   units = "in",
#   dpi = 300
# )
```

**Interpretation:**

- 0 = approximately average relative concern

- Positive = higher relative concern

- Negative = lower relative concern

| Domain | Distribution | What the histogram shows |
|------------------------|------------------------|------------------------|
| **Acute Coronary Syndrome** | **Strongly right-skewed** | Large concentration around low/average scores with a long positive tail and a few extreme high scores. |
| **Stroke** | **Very strongly right-skewed** | Even more concentrated around the lower range, with a pronounced positive tail and extreme observations. |
| **Atrial Fibrillation** | **Right-skewed / irregular** | Most observations are concentrated around the centre/lower range, with a positive tail and an isolated high value. |
| **Cholesterol** | **Approximately symmetric** | Broad, reasonably balanced distribution around zero, although not perfectly normal. |
| **Hypertension** | **Approximately symmetric** | Fairly bell-shaped and centred around zero. |
| **Heart Failure** | **Slightly to moderately left-skewed** | Most observations are around zero, but the negative tail extends further than the positive tail. |
| **Diabetes** | **Moderately left-skewed** | Clear concentration around zero with a longer negative tail and some extreme negative observations. |

```{r}
#| label: boxplot
gp_domain_composite_score |>
  filter(!is.na(domain_standardised_score)) |>
  ggplot(
    aes(
      x = domain,
      y = domain_standardised_score
    )
  ) +
  geom_boxplot() +
  coord_flip() +
  labs(
    title = "Domain score distributions and potential outliers",
    x = NULL,
    y = "Standardised domain score"
  )
```

```{r}
#| label: outliers
outliers <- gp_domain_composite_score |>
  filter(
    domain %in% c(
      "Stroke",
      "Acute Coronory Syndrome",
      "Diabetes"
    ),
    abs(domain_standardised_score) > 3
  ) |>
  arrange(
    domain,
    desc(domain_standardised_score)
  ) |>
  select(
    geography_code,
    domain,
    domain_standardised_score,
    indicators_used,
    indicators_expected,
    completeness
  )

outlier_practices <- outliers |>
  select(
    geography_code,
    domain
  )

outlier_detail <- data |>
  semi_join(
    outlier_practices,
    by = c(
      "geography_code",
      "domain"
    )
  ) |>
  filter(
    indicator_polarity %in% c(1, -1)
  ) |>
  mutate(
    small_denominator = denominator < 30
  ) |>
  select(
    geography_code,
    domain,
    indicator_id,
    indicator_value,
    numerator,
    denominator,
    small_denominator,
    indicator_polarity
  )

knitr::kable(outlier_detail)
```

- Most domain scores are centred around zero, but the distributions vary in their degree of skewness and presence of outliers.

- ACS and Stroke show clear positive outliers and right-skewed distributions, while Diabetes and Heart Failure contain several negative outliers.

- Cholesterol and Hypertension appear comparatively symmetric with no prominent extreme observations.

- Atrial Fibrillation is also relatively compact, although one positive outlier is present.

```{r}
#| label: relationship-between-denominator-size-admission-rates
data |>
  filter(
    trimws(indicator_level) == "GP Level",
    domain %in% c(
      "Stroke",
      "Acute Coronory Syndrome",
      "Diabetes"
    )
  ) |>
  mutate(
    denominator = as.numeric(denominator),
    indicator_value = as.numeric(indicator_value)
  ) |>
  ggplot(
    aes(
      x = denominator,
      y = indicator_value
    )
  ) +
  geom_point(alpha = 0.6) +
  facet_wrap(
    ~ domain,
    scales = "free"
  ) +
  labs(
    title = "Relationship between denominator size and indicator values",
    x = "Denominator",
    y = "Indicator value"
  ) +
  theme_minimal()
```

- **ACS** and **Stroke:** practices with smaller denominators can have much higher admission rates, while many practices have zero events. This is expected for **rare-event events combined with variation in practice population size.**

- No clear patterns for diabetes, which makes sense because it's a percentage based on an eligible diabetic population rather than a rare admission event.

  ACS and Stroke remained right-skewed after correcting the population denominators. Further investigation showed that high rates were predominantly associated with smaller practice populations, reflecting the instability of rare-event rates rather than necessarily data-quality problems.

## Domain correlation

```{r}
#| label: domain-correlation
gp_cluster_complete <- gp_cluster_data |>
  drop_na() # 168, 4% removed

nrow(gp_cluster_data) # 175 
nrow(gp_cluster_complete) # 168

domain_correlation <- gp_cluster_complete |>
  select(-geography_code) |>
  cor(
    use = "pairwise.complete.obs"
  )

domain_correlation 

domain_correlation |>
  as.data.frame() |>
  rownames_to_column("domain_1") |>
  pivot_longer(
    -domain_1,
    names_to = "domain_2",
    values_to = "correlation"
  ) |>
  ggplot(
    aes(
      x = domain_1,
      y = domain_2,
      fill = correlation
    )
  ) +
  scale_fill_gradient2(
    low = "steelblue",
    mid = "white",
    high = "firebrick",
    midpoint = 0,
    limits = c(-1, 1)
  ) +
  geom_tile() +
  geom_text(
    aes(
      label = sprintf("%.1f", correlation)
    ),
    size = 3
  ) +
  coord_equal() +
  labs(
    title = "Correlation between cardiovascular domain scores",
    x = NULL,
    y = NULL,
    fill = "Correlation"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(
      angle = 45,
      hjust = 1
    ),
    panel.grid = element_blank()
  )
```

**No strong multicollinearity was identified between the cardiovascular domain scores. Most correlations were weak to moderate, suggesting that the domains retain distinct information about practice cardiovascular profiles**.

**Positive correlations:**

A positive correlation means higher concern in one domain tends to coincide with higher concern in another.

The strongest association was between **Diabetes and Hypertension (r = 0.5)**, indicating that practices with greater relative concern in diabetes treatment-target achievement also tended to show greater hypertension-related concern. But `0.5` is still not high enough to suggest that the two domains are essentially duplicating one another.

Moderate positive associations were also observed between **Cholesterol and Hypertension (r = 0.4)** and **Diabetes and Cholesterol (r = 0.3)**.

**Negative correlations:**

A negative correlation means higher concern in one domain tends to coincide with lower concern in another.

Mostly very weak negative correlations

```{r}
#| label: spearman
#| output: false


domain_correlation_spearman <- gp_cluster_complete |>
  select(-geography_code) |>
  cor(
    method = "spearman",
    use = "pairwise.complete.obs"
  )

domain_correlation_spearman
```

**Relationship between two variables:**

```{r}
#| label: relationship-between-two-vars
cor_value <- cor(
  gp_cluster_complete$Diabetes,
  gp_cluster_complete$Hypertension,
  use = "complete.obs"
)

gp_cluster_complete |>
  ggplot(
    aes(
      x = Diabetes,
      y = Hypertension
    )
  ) +
  geom_point(
    alpha = 0.6
  ) +
  geom_smooth(
    method = "lm",
    se = FALSE
  ) +
  annotate(
    "text",
    x = Inf,
    y = Inf,
    label = paste0(
      "r = ",
      round(cor_value, 1)
    ),
    hjust = 1.2,
    vjust = 1.5,
    size = 4
  ) +
  labs(
    title = "Diabetes and Hypertension domain concern scores",
    subtitle = "Each point represents a GP practice",
    x = "Diabetes concern score",
    y = "Hypertension concern score"
  ) +
  theme_minimal()
```

```{r}
#| label: scatterplot-matrix

library(GGally)

library(GGally)

cor_coloured <- function(data, mapping, ...) {
  
  x <- eval_data_col(data, mapping$x)
  y <- eval_data_col(data, mapping$y)
  
  test <- cor.test(x, y)
  
  r <- test$estimate
  p <- test$p.value
  
  stars <- case_when(
    p < 0.001 ~ "***",
    p < 0.01  ~ "**",
    p < 0.05  ~ "*",
    TRUE      ~ ""
  )
  
  label <- paste0(
    "Corr: ",
    sprintf("%.1f", r),
    stars
  )
  
  # Highlight statistically significant correlations
  text_colour <- ifelse(
    p < 0.05,
    "firebrick",
    "grey30"
  )
  
  ggplot() +
    annotate(
      "text",
      x = 0.5,
      y = 0.5,
      label = label,
      colour = text_colour,
      size = 3
    ) +
    theme_void()
}

gp_cluster_complete |>
  rename(
    ACS = `Acute Coronory Syndrome`,
    AF  = `Atrial Fibrillation`,
    HF  = `Heart Failure`
  ) |>
  select(-geography_code) |>
  ggpairs(
    lower = list(
      continuous = wrap(
        "points",
        alpha = 0.4,
        size = 0.8
      )
    ),
    upper = list(
      continuous = cor_coloured
    ),
    diag = list(
      continuous = "densityDiag"
    )
  )
```

## PCA analysis

```{r}
#| label: pca-analysis
gp_matrix <- gp_cluster_complete |>
  select(-geography_code) |>
  as.matrix()

gp_pca <- prcomp(
  gp_matrix,
  center = TRUE,
  scale. = TRUE
)

pca_scores <- as_tibble(
  gp_pca$x[, 1:2]
) |>
  mutate(
    geography_code = gp_cluster_complete$geography_code
  )

variance_explained <- 100 * gp_pca$sdev^2 / sum(gp_pca$sdev^2)

pc1_var <- variance_explained[1]
pc2_var <- variance_explained[2]

ggplot(
  pca_scores,
  aes(
    x = PC1,
    y = PC2
  )
) +
  geom_point() +
  labs(
    title = "PCA exploration of GP cardiovascular profiles",
    x = paste0(
      "Principal component 1 (",
      round(pc1_var, 1),
      "%)"
    ),
    y = paste0(
      "Principal component 2 (",
      round(pc2_var, 1),
      "%)"
    )
  ) +
  theme_minimal()
```

**There are no obvious discrete groups**.

**The first two principal components showed a continuous distribution of GP practices with no visually distinct groups. A small number of practices had more unusual profiles at the extremes of the PCA space, but there was no clear evidence of naturally separated clusters**.

Most practices lie roughly around PC1 from about −2.5 to +2.5 and PC2 from about −2 to +2. There are then several practices extending strongly to the right.plus a few vertically extreme points. That suggests some practices have unusual multidomain profiles, but there are still no obvious, cleanly separated clusters.

- **PCA identified a multidimensional structure in cardiovascular practice profiles. PC1 (26.7%) was predominantly associated with Diabetes, Hypertension and Cholesterol concern, while PC2 (17.7%) represented a contrast involving Stroke/AF and ACS/Hypertension/HF.**

- **The first two components explained 44.5% of total variation, increasing to 73.7% across four components.**

- **Practices formed a continuous distribution in PCA space rather than clearly separated groups, providing limited visual evidence of natural clusters.**

**PCA loadings:**

PC1 explains **26.7%** of the variation.

**PC1 — treatment/risk-factor management dimension**

```{r}
#| label: pca-loadings
pca_loadings <- gp_pca$rotation |>
  as.data.frame() |>
  rownames_to_column("domain") |>
  as_tibble()

pca_loadings
```

The dominant loadings in PC1 are:

| Domain        | PC1        |
|---------------|------------|
| Diabetes      | **+0.571** |
| Hypertension  | **+0.521** |
| Cholesterol   | **+0.518** |
| Heart Failure | −0.274     |
| AF            | +0.195     |
| Stroke        | +0.148     |
| ACS           | −0.004     |

PC2 explains another **17.7%**.

**PC2 — acute/event and AF profile**

| Domain        | PC2        |
|---------------|------------|
| Stroke        | **+0.542** |
| AF            | **+0.469** |
| ACS           | **−0.431** |
| Hypertension  | **−0.422** |
| Heart Failure | −0.312     |
| Diabetes      | −0.139     |
| Cholesterol   | +0.079     |

**PCA variance:**

The first two PCs explain only **44.5%**. Therefore, the scatterplot represents **less than half of the total information in the 7 domain scores.** We need 4 PCs to represent about 74% of the variation.

This indicate practices differ across **multiple dimensions**, rather than there being one or two simple dimensions that explain everything.

```{r}
#| label: pca-variance
pca_variance <- tibble(
  component = paste0("PC", seq_along(gp_pca$sdev)),
  variance_explained =
    100 * gp_pca$sdev^2 / sum(gp_pca$sdev^2)
) |>
  mutate(
    cumulative_variance = cumsum(variance_explained)
  )

pca_variance
```

```{r}
#| label: scree-plot
# install.packages("factoextra")
library(factoextra)

fviz_eig(
  gp_pca,
  addlabels = TRUE,
  ylim = c(0, 35)
) +
  labs(
    title = "Variance explained by principal components",
    x = "Principal component",
    y = "Variance explained (%)"
  )
```

```{r}
#| label: cumulative-var-plot
pca_variance |>
  ggplot(
    aes(
      x = component,
      y = cumulative_variance,
      group = 1
    )
  ) +
  geom_line() +
  geom_point() +
  geom_text(
    aes(
      label = paste0(round(cumulative_variance, 1), "%")
    ),
    vjust = -0.7
  ) +
  geom_hline(
    yintercept = c(70, 80),
    linetype = "dashed"
  ) +
  scale_y_continuous(
    limits = c(0, 105),
    breaks = seq(0, 100, 10)
  ) +
  labs(
    title = "Cumulative variance explained",
    x = "Principal component",
    y = "Cumulative variance explained (%)"
  ) +
  theme_minimal()
```

The first two principal components captured only 45% of total variation, while four components were required to explain approximately 74%. This indicated that cardiovascular practice profiles were multidimensional and could not be adequately represented by a single underlying performance dimension. Therefore, the individual cardiovascular domain scores were retained for clustering to preserve clinically meaningful differences between practices.

# Clustering algorithms

| Method | Why include it |
|------------------------------------|------------------------------------|
| **K-means** | Simple, widely understood, good baseline; but sensitive to outliers and assumes roughly compact/spherical clusters |
| **PAM (K-medoids)** | More robust to extreme observations because cluster centres are actual practices rather than means |
| **Hierarchical clustering** | Doesn't require fixing `k` before fitting and lets us explore nested practice groupings/dendrograms |

```{r}

gp_codes <- gp_cluster_complete$geography_code

gp_matrix <- gp_cluster_complete |>
  select(-geography_code) |>
  as.matrix()

gp_matrix_scaled <- scale(gp_matrix)

k_values <- 2:8
```

## K-means

```{r}
library(cluster)
library(tidyverse)

set.seed(123)

kmeans_results <- map_dfr(
  k_values,
  function(k) {
    
    model <- kmeans(
      gp_matrix_scaled,
      centers = k,
      nstart = 100,
      iter.max = 100
    )
    
    sil <- silhouette(
      model$cluster,
      dist(gp_matrix_scaled)
    )
    
    tibble(
      algorithm = "K-means",
      k = k,
      average_silhouette = mean(sil[, "sil_width"])
    )
  }
)
```

## PAM

```{r}
pam_results <- map_dfr(
  k_values,
  function(k) {
    
    model <- cluster::pam(
      gp_matrix_scaled,
      k = k,
      metric = "euclidean"
    )
    
    tibble(
      algorithm = "PAM",
      k = k,
      average_silhouette =
        model$silinfo$avg.width
    )
  }
)
```

## Hierarchical clustering

```{r}
gp_distance <- dist(
  gp_matrix_scaled,
  method = "euclidean"
)

hierarchical_model <- hclust(
  gp_distance,
  method = "ward.D2"
)

hierarchical_results <- map_dfr(
  k_values,
  function(k) {
    
    clusters <- cutree(
      hierarchical_model,
      k = k
    )
    
    sil <- silhouette(
      clusters,
      gp_distance
    )
    
    tibble(
      algorithm = "Hierarchical",
      k = k,
      average_silhouette =
        mean(sil[, "sil_width"])
    )
  }
)
```

## Compare all three

```{r}
algorithm_comparison <- bind_rows(
  kmeans_results,
  pam_results,
  hierarchical_results
)

knitr::kable(algorithm_comparison)
```

### Cluster separation

**Average silhouette width** tells us **how well the observations fit within their assigned clusters and how well separated the clusters are from each other.**

For each GP practice, the silhouette score compares:

- **Cohesion:** how close that practice is to other practices in its own cluster.

- **Separation:** how far it is from practices in the nearest alternative cluster.

The silhouette score ranges from -1 to +1:

| Silhouette | Meaning |
|------------------------------------|------------------------------------|
| Close to **+1** | Practice fits its cluster very well and is far from other clusters |
| Around **0** | Practice lies between clusters; clusters overlap |
| Below **0** | Practice may fit another cluster better |

Which combination of algorithm and number of clusters produces the most cohesive and well-separated groups of GP practices?

```{r}

algorithm_comparison |>
  ggplot(
    aes(
      x = k,
      y = average_silhouette,
      colour = algorithm,
      linetype = algorithm,
      group = algorithm
    )
  ) +
  geom_line(linewidth = 1) +
  geom_point(size = 2.5) +
  scale_x_continuous(
    breaks = k_values
  ) +
  labs(
    title = "Comparison of clustering algorithms",
    x = "Number of clusters",
    y = "Average silhouette width",
    colour = "Algorithm",
    linetype = "Algorithm"
  ) +
  theme_minimal()
```

K-means with k = 4 achieved the highest average silhouette width, but the relatively low value (\~0.18) indicated weak cluster separation, suggesting cardiovascular practice profiles may vary more continuously rather than forming strongly defined natural groups.

### Cluster size

```{r}
get_cluster_sizes <- function(cluster_vector, algorithm, k) {
  
  tibble(
    geography_code = gp_codes,
    cluster = cluster_vector
  ) |>
    count(cluster) |>
    mutate(
      algorithm = algorithm,
      k = k,
      percentage = 100 * n / sum(n)
    )
}

set.seed(123)

kmeans_4 <- kmeans(
  gp_matrix_scaled,
  centers = 4,
  nstart = 100,
  iter.max = 100
)

pam_4 <- pam(
  gp_matrix_scaled,
  k = 4,
  metric = "euclidean"
)

hierarchical_4 <- cutree(
  hierarchical_model,
  k = 4
)

cluster_sizes_4 <- bind_rows(
  
  get_cluster_sizes(
    kmeans_4$cluster,
    "K-means",
    4
  ),
  
  get_cluster_sizes(
    pam_4$clustering,
    "PAM",
    4
  ),
  
  get_cluster_sizes(
    hierarchical_4,
    "Hierarchical",
    4
  )
)



cluster_sizes_4 |>
  ggplot(
    aes(
      x = factor(cluster),
      y = n,
      fill = factor(cluster)
    )
  ) +
  geom_col() +
  geom_text(
    aes(
      label = paste0(
        n,
        " (",
        round(percentage, 0),
        "%)"
      )
    ),
    vjust = -0.4
  ) +
  facet_wrap(
    ~ algorithm
  ) +
  labs(
    title = "Cluster sizes for four-cluster solutions",
    x = "Cluster",
    y = "Number of GP practices",
    fill = "Cluster"
  ) +
  theme_minimal() +
  theme(
    legend.position = "none"
  )
```

At `k = 4`:

The three algorithms are finding **very different structures** in the same data.

For k-means:

- Despite having the highest silhouette score (\~0.18), it is essentially identifying two large groups plus two relatively small groups. The latter may not necessarily be wrong, but they could represent genuinely unusual cardiovascular profiles, but they warrant further investigation.

For hierarchical clustering:

- Behaves similarly to k-means - 2 large clusters and 2 smaller ones. Given that two different algorithms are finding broadly similar size *structures,* it gives some support to the possibility that there are similar atypical practice profiles.

For PAM:

- Behaves quite different from the other two algorithms: there are 3 clusters which are much more balanced with one very small cluster.

### Cluster shape

```{r}
# k-means ------------------
fviz_cluster(
  kmeans_4,
  data = gp_matrix_scaled,
  palette = "jco",          # colourblind-safe journal palette
  ellipse.type = "t",       # t-distribution concentration ellipse
  repel = FALSE,             # keep state labels from overlapping
  ggtheme = theme_minimal()
)
# Hierarchical ---------------
fviz_dend(
  hierarchical_model,
  k = 4,
  cex = 0.3,
  rect = TRUE,
  rect_fill = TRUE,
  show_labels = FALSE,
  color_labels_by_k = TRUE
)
# PAM -----------------------
fviz_cluster(
  pam_4,
  palette = "jco",          # colourblind-safe journal palette
  ellipse.type = "t",       # t-distribution concentration ellipse
  repel = FALSE,             # keep state labels from overlapping
  ggtheme = theme_minimal()
)
```

### Cluster profiles

Comparing cluster profiles for all 3 algorithms at k = 4 to see what each algorithm is actually separating.

```{r}
compare_cluster_profiles <- function(data, cluster_vector, algorithm) {
  
  data |>
    mutate(
      cluster = factor(cluster_vector)
    ) |>
    pivot_longer(
      cols = -c(geography_code, cluster),
      names_to = "domain",
      values_to = "score"
    ) |>
    group_by(
      cluster,
      domain
    ) |>
    summarise(
      mean_score = mean(score, na.rm = TRUE),
      .groups = "drop"
    ) |>
    mutate(
      algorithm = algorithm
    )
}

cluster_profiles_4 <- bind_rows(
  compare_cluster_profiles(
    gp_cluster_complete,
    hierarchical_4,
    "Hierarchical"
  ),
  compare_cluster_profiles(
    gp_cluster_complete,
    kmeans_4$cluster,
    "K-means"
  ),
  compare_cluster_profiles(
    gp_cluster_complete,
    pam_4$clustering,
    "PAM"
  )
)

cluster_profiles_4 |>
  ggplot(
    aes(
      x = domain,
      y = mean_score,
      group = cluster,
      colour = cluster
    )
  ) +
  geom_hline(
    yintercept = 0,
    linetype = "dashed"
  ) +
  geom_line() +
  geom_point() +
  facet_wrap(
    ~ algorithm
  ) +
  labs(
    title = "Cardiovascular profiles of four-cluster solutions",
    x = "Clinical domain",
    y = "Mean standardised concern score",
    colour = "Cluster"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(
      angle = 45,
      hjust = 1
    )
  )
```

```{r}
get_cluster_profiles <- function(
    data,
    cluster_vector,
    algorithm
) {
  
  data |>
    mutate(
      cluster = factor(cluster_vector)
    ) |>
    pivot_longer(
      cols = -c(geography_code, cluster),
      names_to = "domain",
      values_to = "score"
    ) |>
    group_by(
      cluster,
      domain
    ) |>
    summarise(
      mean_score = mean(score, na.rm = TRUE),
      .groups = "drop"
    ) |>
    mutate(
      algorithm = algorithm
    )
}

cluster_profiles_4 <- bind_rows(
  
  get_cluster_profiles(
    gp_cluster_complete,
    kmeans_4$cluster,
    "K-means"
  ),
  
  get_cluster_profiles(
    gp_cluster_complete,
    pam_4$clustering,
    "PAM"
  ),
  
  get_cluster_profiles(
    gp_cluster_complete,
    hierarchical_4,
    "Hierarchical"
  )
)

cluster_profiles_band_4 <- cluster_profiles_4 |>
  mutate(
    concern_band = case_when(
      mean_score >= 1    ~ "Higher relative concern",
      mean_score >= 0.5  ~ "Moderate relative concern",
      mean_score > -0.5  ~ "Around average",
      mean_score > -1    ~ "Lower relative concern",
      TRUE               ~ "Very low relative concern"
    ),
    
    concern_band = factor(
      concern_band,
      levels = c(
        "Very low relative concern",
        "Lower relative concern",
        "Around average",
        "Moderate relative concern",
        "Higher relative concern"
      )
    )
  )

profile_annotations <- tibble(
  algorithm = c(
    "Hierarchical", "Hierarchical",
    "K-means", "K-means",
    "PAM", "PAM"
  ),
  cluster = factor(
    c(3, 4, 4, 3, 4, 3),
    levels = 1:4
  ),
  domain = "Profile",
  profile = c(
    "Higher cardiometabolic concern",
    "Lower cardiometabolic concern",
    "Higher cardiometabolic concern",
    "Lower cardiometabolic concern",
    "Higher cardiometabolic concern",
    "Lower cardiometabolic concern"
  )
)

domain_order <- c(
  "Acute Coronory Syndrome",
  "Atrial Fibrillation",
  "Cholesterol",
  "Diabetes",
  "Heart Failure",
  "Hypertension",
  "Stroke",
  "Profile"
)

cluster_profiles_band_4 <- cluster_profiles_band_4 |>
  mutate(
    domain = factor(
      domain,
      levels = domain_order
    )
  )

profile_annotations <- profile_annotations |>
  mutate(
    domain = factor(
      domain,
      levels = domain_order
    )
  )

cluster_profiles_band_4 |>
  ggplot(
    aes(
      x = domain,
      y = cluster,
      fill = concern_band
    )
  ) +
  geom_tile(
    colour = "white",
    linewidth = 1
  ) +
  geom_text(
    aes(
      label = round(mean_score, 1)
    ),
    size = 3
  ) +
  
  # Add profile interpretation
  geom_text(
    data = profile_annotations,
    aes(
      x = domain,
      y = cluster,
      label = profile
    ),
    inherit.aes = FALSE,
    hjust = 0,
    size = 3.2
  ) +
  
  facet_wrap(
    ~ algorithm,
    ncol = 1
  ) +
  scale_x_discrete(
    drop = FALSE,
    labels = c(
      "Profile" = ""
    )
  ) +
  scale_fill_manual(
    values = c(
      "Very low relative concern" = "#4575B4",
      "Lower relative concern" = "#91BFDB",
      "Around average" = "#FFFFBF",
      "Moderate relative concern" = "#FC8D59",
      "Higher relative concern" = "#D73027"
    ),
    drop = FALSE
  ) +
  labs(
    title = "Cardiovascular concern profiles by clustering algorithm",
    subtitle = "Four-cluster solutions",
    x = "Clinical domain",
    y = "Practice cluster",
    fill = "Relative concern"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(
      angle = 45,
      hjust = 1
    ),
    panel.grid = element_blank()
  )
```

**Summary:**

1.  A small **very high Stroke** group

2.  A small/high **ACS** group, particularly with K-means and hierarchical

3.  A larger **lower cardiometabolic concern profile**

4.  A larger **higher cardiometabolic concern profile**

Across clustering algorithms, some consistent cardiovascular profiles emerged.

In particular, larger groups were differentiated by higher or lower concern across the correlated Diabetes, Cholesterol and Hypertension domains, consistent with the structure identified by PCA.

Smaller clusters were characterized by extreme Stroke or ACS scores.

However, overall silhouette widths remained low, suggesting substantial overlap between practices and weak evidence for four clearly separated natural clusters.

**Using median score for clusters profiling**

```{r}
get_cluster_profiles <- function(
    data,
    cluster_vector,
    algorithm
) {
  
  data |>
    mutate(
      cluster = factor(cluster_vector)
    ) |>
    pivot_longer(
      cols = -c(geography_code, cluster),
      names_to = "domain",
      values_to = "score"
    ) |>
    group_by(
      cluster,
      domain
    ) |>
    summarise(
      median_score = median(score, na.rm = TRUE),
      .groups = "drop"
    ) |>
    mutate(
      algorithm = algorithm
    )
}

cluster_profiles_4 <- bind_rows(
  
  get_cluster_profiles(
    gp_cluster_complete,
    kmeans_4$cluster,
    "K-means"
  ),
  
  get_cluster_profiles(
    gp_cluster_complete,
    pam_4$clustering,
    "PAM"
  ),
  
  get_cluster_profiles(
    gp_cluster_complete,
    hierarchical_4,
    "Hierarchical"
  )
)

cluster_profiles_band_4 <- cluster_profiles_4 |>
  mutate(
    concern_band = case_when(
      median_score >= 1    ~ "Higher relative concern",
      median_score >= 0.5  ~ "Moderate relative concern",
      median_score > -0.5  ~ "Around average",
      median_score > -1    ~ "Lower relative concern",
      TRUE               ~ "Very low relative concern"
    ),
    
    concern_band = factor(
      concern_band,
      levels = c(
        "Very low relative concern",
        "Lower relative concern",
        "Around average",
        "Moderate relative concern",
        "Higher relative concern"
      )
    )
  )

profile_annotations <- tibble(
  algorithm = c(
    "Hierarchical", "Hierarchical",
    "K-means", "K-means",
    "PAM", "PAM"
  ),
  cluster = factor(
    c(3, 4, 4, 3, 4, 3),
    levels = 1:4
  ),
  domain = "Profile",
  profile = c(
    "Higher cardiometabolic concern",
    "Lower cardiometabolic concern",
    "Higher cardiometabolic concern",
    "Lower cardiometabolic concern",
    "Higher cardiometabolic concern",
    "Lower cardiometabolic concern"
  )
)

domain_order <- c(
  "Acute Coronory Syndrome",
  "Atrial Fibrillation",
  "Cholesterol",
  "Diabetes",
  "Heart Failure",
  "Hypertension",
  "Stroke",
  "Profile"
)

cluster_profiles_band_4 <- cluster_profiles_band_4 |>
  mutate(
    domain = factor(
      domain,
      levels = domain_order
    )
  )

profile_annotations <- profile_annotations |>
  mutate(
    domain = factor(
      domain,
      levels = domain_order
    )
  )

cluster_profiles_band_4 |>
  ggplot(
    aes(
      x = domain,
      y = cluster,
      fill = concern_band
    )
  ) +
  geom_tile(
    colour = "white",
    linewidth = 1
  ) +
  geom_text(
    aes(
      label = round(median_score, 1)
    ),
    size = 3
  ) +
  
  # Add profile interpretation
  geom_text(
    data = profile_annotations,
    aes(
      x = domain,
      y = cluster,
      label = profile
    ),
    inherit.aes = FALSE,
    hjust = 0,
    size = 3.2
  ) +
  
  facet_wrap(
    ~ algorithm,
    ncol = 1
  ) +
  scale_x_discrete(
    drop = FALSE,
    labels = c(
      "Profile" = ""
    )
  ) +
  scale_fill_manual(
    values = c(
      "Very low relative concern" = "#4575B4",
      "Lower relative concern" = "#91BFDB",
      "Around average" = "#FFFFBF",
      "Moderate relative concern" = "#FC8D59",
      "Higher relative concern" = "#D73027"
    ),
    drop = FALSE
  ) +
  labs(
    title = "Cardiovascular concern profiles by clustering algorithm",
    subtitle = "Four-cluster solutions",
    x = "Clinical domain",
    y = "Practice cluster",
    fill = "Relative concern"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(
      angle = 45,
      hjust = 1
    ),
    panel.grid = element_blank()
  )
```

## Sensitivity analysis

Repeat the clustering without ACS and Stroke

```{r}
# Sensitivity dataset: remove ACS and Stroke -------------------------------

gp_matrix_sensitivity <- gp_cluster_complete |>
  select(
    -geography_code,
    -`Acute Coronory Syndrome`,
    -Stroke
  ) |>
  scale()

k_values <- 2:8

# K-means -----------------------------------------------------------------

set.seed(123)

kmeans_results_sensitivity <- map_dfr(
  k_values,
  function(k) {
    
    model <- kmeans(
      gp_matrix_sensitivity,
      centers = k,
      nstart = 100,
      iter.max = 100
    )
    
    sil <- silhouette(
      model$cluster,
      dist(gp_matrix_sensitivity)
    )
    
    tibble(
      algorithm = "K-means",
      k = k,
      average_silhouette = mean(sil[, "sil_width"])
    )
  }
)

# PAM ---------------------------------------------------------------------

pam_results_sensitivity <- map_dfr(
  k_values,
  function(k) {
    
    model <- cluster::pam(
      gp_matrix_sensitivity,
      k = k,
      metric = "euclidean"
    )
    
    tibble(
      algorithm = "PAM",
      k = k,
      average_silhouette = model$silinfo$avg.width
    )
  }
)

# Hierarchical clustering -------------------------------------------------

gp_distance_sensitivity <- dist(
  gp_matrix_sensitivity,
  method = "euclidean"
)

hierarchical_model_sensitivity <- hclust(
  gp_distance_sensitivity,
  method = "ward.D2"
)

hierarchical_results_sensitivity <- map_dfr(
  k_values,
  function(k) {
    
    clusters <- cutree(
      hierarchical_model_sensitivity,
      k = k
    )
    
    sil <- silhouette(
      clusters,
      gp_distance_sensitivity
    )
    
    tibble(
      algorithm = "Hierarchical",
      k = k,
      average_silhouette = mean(sil[, "sil_width"])
    )
  }
)

# Combine results ---------------------------------------------------------

algorithm_comparison_sensitivity <- bind_rows(
  kmeans_results_sensitivity,
  pam_results_sensitivity,
  hierarchical_results_sensitivity
)

# Plot --------------------------------------------------------------------

algorithm_comparison_sensitivity |>
  ggplot(
    aes(
      x = k,
      y = average_silhouette,
      colour = algorithm,
      linetype = algorithm,
      group = algorithm
    )
  ) +
  geom_line(linewidth = 1) +
  geom_point(size = 2.5) +
  scale_x_continuous(
    breaks = k_values
  ) +
  labs(
    title = "Sensitivity analysis: clustering without ACS and Stroke",
    subtitle = "Comparison of clustering algorithms",
    x = "Number of clusters",
    y = "Average silhouette width",
    colour = "Algorithm",
    linetype = "Algorithm"
  ) +
  theme_minimal()
```

Sensitivity analysis showed that the apparent strong two-cluster solution was largely driven by rare-event ACS and Stroke measures. After excluding these domains, cluster separation was weak across all algorithms, suggesting cardiovascular need may be better represented as a continuum of practice profiles rather than discrete groups.

Plot clusters on the PCA space:

```{r}
set.seed(123)

kmeans_5 <- kmeans(
  gp_matrix_sensitivity,
  centers = 5,
  nstart = 100
)

# Run PCA on the same sensitivity matrix
pca_sensitivity <- prcomp(
  gp_matrix_sensitivity,
  center = FALSE,
  scale. = FALSE
)

pca_cluster_plot <- as_tibble(
  pca_sensitivity$x[, 1:2]
) |>
  mutate(
    geography_code = gp_cluster_complete$geography_code,
    cluster = factor(kmeans_5$cluster)
  )

pca_cluster_plot |>
  ggplot(
    aes(
      x = PC1,
      y = PC2,
      colour = cluster
    )
  ) +
  geom_point(
    size = 2.5,
    alpha = 0.8
  ) +
  labs(
    title = "GP cardiovascular practice clusters",
    subtitle = "K-means, k = 5; ACS and Stroke excluded",
    x = "Principal component 1",
    y = "Principal component 2",
    colour = "Cluster"
  ) +
  theme_minimal()
```

```{r}
set.seed(123)
km5.res <- kmeans(gp_matrix_sensitivity, centers = 5, nstart = 100)


fviz_cluster(
  km5.res, data = gp_matrix_sensitivity,
  palette = "jco",            # colourblind-safe journal palette
  ellipse.type = "euclid",    # concentration ellipse around each cluster
  star.plot = FALSE,           # segments from each centroid to its members
  repel = FALSE,               # keep state labels from overlapping
  ggtheme = theme_minimal(),
  show.clust.cent = TRUE,
  geom = "point"
)+
  labs(
    title = "K-means clustering of GP cardiovascular profiles",
    subtitle = "Five-cluster solution; ACS and Stroke excluded"
  )
```

With ACS and Stroke removed, the 5 clusters exist mathematically, but several are not strongly separated.

The axes `Dim1 (31.6%)` and `Dim2 (24.8%)` are the first two principal dimensions used by `fviz_cluster()` to project a five-dimensional data into 2D. Together they display about **56.4% of the variation**, so remember that some separation may exist in dimensions not visible here.

**Cluster 2 (yellow)** is the clearest cluster. It sits predominantly on the right-hand side and has relatively little overlap with the others. This suggests those practices have a genuinely different cardiovascular profile on the dimensions driving this projection.

**Cluster 1 (dark blue)** also has some structure, mainly in the lower-left region. However, its boundary overlaps with clusters 3 and 5.

**Clusters 3, 4 and 5** overlap substantially around the centre/upper-left area. Their ellipses intersect considerably, and many observations are intermixed.

Plot the cluster profiles:

```{r}
cluster_profiles_5 <- gp_cluster_complete |>
  select(
    geography_code,
    `Atrial Fibrillation`,
    Cholesterol,
    Diabetes,
    `Heart Failure`,
    Hypertension
  ) |>
  mutate(
    cluster = factor(km5.res$cluster)
  ) |>
  pivot_longer(
    cols = c(
      `Atrial Fibrillation`,
      Cholesterol,
      Diabetes,
      `Heart Failure`,
      Hypertension
    ),
    names_to = "domain",
    values_to = "score"
  ) |>
  group_by(
    cluster,
    domain
  ) |>
  summarise(
    mean_score = mean(score, na.rm = TRUE),
    .groups = "drop"
  )

cluster_profiles_5 |>
  ggplot(
    aes(
      x = domain,
      y = mean_score,
      colour = cluster,
      group = cluster
    )
  ) +
  geom_hline(
    yintercept = 0,
    linetype = "dashed"
  ) +
  geom_line(
    linewidth = 1
  ) +
  geom_point(
    size = 2.5
  ) +
  labs(
    title = "Cardiovascular profile by practice cluster",
    x = "Clinical domain",
    y = "Mean standardised concern score",
    colour = "Cluster"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(
      angle = 45,
      hjust = 1
    )
  )
```

A five-cluster K-means solution produced clinically interpretable cardiovascular profiles, but overall cluster separation was weak, indicating that the groups should be treated as exploratory segments rather than definitive classifications.

```{r}
cluster_profiles_band <- cluster_profiles_5 |>
  mutate(
    concern_band = case_when(
      mean_score >= 1    ~ "Higher relative concern",
      mean_score >= 0.5  ~ "Moderate relative concern",
      mean_score > -0.5  ~ "Around average",
      mean_score > -1    ~ "Lower relative concern",
      TRUE               ~ "Very low relative concern"
    ),
    concern_band = factor(
      concern_band,
      levels = c(
        "Very low relative concern",
        "Lower relative concern",
        "Around average",
        "Moderate relative concern",
        "Higher relative concern"
      )
    )
  )

cluster_profiles_band |>
  ggplot(
    aes(
      x = domain,
      y = cluster,
      fill = concern_band
    )
  ) +
  geom_tile(
    colour = "white",
    linewidth = 1
  ) +
  geom_text(
    aes(
      label = round(mean_score, 1)
    )
  ) +
  scale_fill_manual(
    values = c(
      "Very low relative concern" = "#4575B4",
      "Lower relative concern" = "#91BFDB",
      "Around average" = "#FFFFBF",
      "Moderate relative concern" = "#FC8D59",
      "Higher relative concern" = "#D73027"
    ),
    drop = FALSE
  ) +
  labs(
    title = "Cardiovascular concern profile by practice cluster",
    x = "Clinical domain",
    y = "Practice cluster",
    fill = "Relative concern"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(
      angle = 45,
      hjust = 1
    ),
    panel.grid = element_blank()
  )
```

0 is approximately average relative to other practices; positive scores indicate greater relative concern; negative scores indicate lower relative concern.

| Cluster | Main interpretation |
|------------------------------------|------------------------------------|
| **1** | **Cholesterol / AF profile** — higher concern for cholesterol (+0.9), AF (+0.6) and HF (+0.5), while diabetes is around average. |
| **2** | **Heart Failure profile** — HF is particularly high (+1.0), while all other domains are substantially below average, especially diabetes (−1.9) and hypertension (−1.2). |
| **3** | **Hypertension profile** — hypertension is the main concern (+0.7), with moderately elevated diabetes and HF (+0.4 each). AF is relatively low (−0.8). |
| **4** | **AF / Diabetes profile** — moderately elevated AF and diabetes (+0.4), while cholesterol (−0.8) and hypertension (−0.7) are relatively low. |
| **5** | **Cholesterol / Diabetes profile** — cholesterol is the strongest concern (+0.7), with some diabetes (+0.4) and hypertension (+0.3), but HF is very low (−1.5). |

K-means identified five potentially interpretable cardiovascular profiles; however, relatively low silhouette scores indicate substantial overlap between groups, so these should be considered exploratory profiles rather than definitive practice classifications.

# Conclusions

- **No strong evidence of naturally distinct practice clusters** was identified across the broader cardiovascular profiles.

<!-- -->

- **Cluster separation was weak across K-means, PAM and hierarchical clustering**, suggesting substantial overlap between practice profiles.

<!-- -->

- Nevertheless, clustering identified **potentially interpretable cardiovascular profiles**, such as practices with relatively greater concern around hypertension, diabetes or cholesterol, but these should be treated as **exploratory rather than definitive classifications**.

<!-- -->

- For risk stratification, **continuous domain concern scores may therefore be more appropriate than forcing practices into discrete clusters**, allowing commissioners to identify practices experiencing challenges across multiple cardiovascular areas and target support accordingly.

<!-- -->

- **ACS and Stroke remain clinically important**, but rare-event measures may require alternative handling, such as longer time periods, pooled data or minimum-denominator/event thresholds, before inclusion in practice-level segmentation.

**Recommendation:** Use the standardised domain concern scores as the primary risk-stratification framework, with clustering retained as an exploratory profiling tool rather than the sole basis for prioritising practices.

# Strengths:

- **Multiple cardiovascular domains and data sources.** Clustering practices was not done using a single performance measure. The analysis combines 7 clinically relevant domains (ACS, AF, Cholesterol, Diabetes, HF, Hypertension and Stroke) built from multiple indicators. This gives a more multidimensional view of practice cardiovascular profiles.

- **Consistent polarity and standardisation across indicators**

- **Systematic assessment of missingness, outliers and rare-event measures**

- **PCA and correlation analysis used to understand underlying structure**

- **Multiple clustering algorithms and values of `k` compared**

- **Cluster selection considered separation, size and clinical interpretability**

# Limitations:

- Weak overall cluster separation (maximum silhouette \~ 0.18)

- ACS/Stroke are skewed rare-event measures sensitive to small populations

- Unequal numbers of indicators contribute to individual domains

- Composite scores may mask within-domain differences

- Cross-sectional analysis, cluster stability over time not established

- Practice profiles may reflect population/case-mix differences as well as performance

# References:

## Indicator list

| indicator_id | indicator_short_name | indicator_description | value_type | value_multiplier | indicator_polarity | domain |
|---:|---|---|---|---:|---:|---|
| 1 | Acute coronary syndrome admission rate in BSol population | Acute coronary syndrome admission rate per 100,000 population in Bsol | Rate | 100000 | -1 | Acute Coronory Syndrome |
| 2 | Atrial fibrillation and anticoagulated | Percentage of patients aged 18 and over with GP recorded atrial fibrillation (AF) and a record of a CHADS2 or CHA2DS2-VASc score of 2 or more, who have received a prescription for any oral anticoagulation drug therapy in the preceding 6 months. | Percentage | 100 | 1 | Atrial Fibrillation |
| 4 | Diabetes – meeting all 3 treatment targets | Percentage of patients meeting all 3 treatment targets for diabetes | Percentage | 100 | 1 | Diabetes |
| 8 | Patients with CVD treated to target cholesterol threshold | Percentage of patients with GP recorded CVD (narrow definition), whose most recent blood cholesterol level is LDL-cholesterol less than or equal to 2.0 mmol/l or non-HDL cholesterol less than or equal to 2.6 mmol/l, in the preceding 12 months | Percentage | 100 | 1 | Cholesterol |
| 9 | Patients with no CVD and a QRISK score >20% treated with a lipid lowering drug | Percentage of atients with no GP recorded CVD and a GP recorded QRISK score of 20% or more, who are currently treated with lipid lowering therapy | Percentage | 100 | 1 | Cholesterol |
| 10 | Patients with GP recorded hypertension | Percentage of patients aged 18 and over with GP recorded hypertension, who have had a blood pressure reading within the preceding 12 months | Percentage | 100 | 1 | Hypertension |
| 11 | Patients at risk of hypertension | Percentage of patients aged 18 and over, whose latest blood pressure value is in the at risk range for hypertension with no GP recorded hypertension | Percentage | 100 | -1 | Hypertension |
| 12 | Patients treated to appropriate threshold (all ages) | Percentage of patients aged 18 and over, with GP recorded hypertension, in whom the last blood pressure reading (measured in the preceding 12 months) is below the age appropriate treatment threshold | Percentage | 100 | 1 | Hypertension |
| 13 | Patients treated to appropriate threshold (age < 80) | Percentage of patients aged 79 and under controlled to age specific blood pressure target | Percentage | 100 | 1 | Hypertension |
| 14 | Patients treated to appropriate threshold (age 80 and over) | Percentage of patients aged 80 and over controlled to age specific blood pressure target | Percentage | 100 | 1 | Hypertension |
| 15 | Patients with CKD treated with LLT | Percentage of patients with GP recorded chronic kidney disease (G3a to G5), who are currently treated with lipid lowering therapy. | Percentage | 100 | 1 | Cholesterol |
| 17 | Stroke admissions rate in BSol registered population | Stroke admission rate per 100,000 population registered at BSol practices | Rate | 100000 | -1 | Stroke |
| 18 | Stroke admissions rate in BSol hospitals | Stroke admissions rate per 100,000 population for admissions to Bsol hospitals | Rate | 100000 | -1 | Stroke |
| 19 | Heart failure admissions rate in Bsol registered population | Heart failure admissions rate per 100,000 population registered at Bsol practices | Rate | 100000 | -1 | Heart Failure |
| 20 | Patients prescribed with RAAS drugs | Percentage of patients prescribed RAAS drug | Percentage | 100 | 1 | Heart Failure |
| 21 | Patients prescribed with Beta-blockers drugs | Percentage of patients prescribed Beta-blockers drugs | Percentage | 100 | 1 | Heart Failure |
| 22 | Patients prescribed with SGLT2 Inhibitor drugs | Percentage of patients prescribed SGLT2 Inhibitor drugs | Percentage | 100 | 1 | Heart Failure |
| 23 | Patients prescribed with RAAS/Beta-blockers/SGLT2 Inhibitor drugs | Percentage of patients prescribed with one of the following drugs: RAAS, Beta-blockers, SGLT2 Inhibitor drugs | Percentage | 100 | 1 | Heart Failure |
| 24 | Patients prescribed with RAAS, Beta-blockers and SGLT2 Inhibitor drugs | Percentage of patients prescribed with all of the following drugs: RAAS, Beta-blockers and SGLT2 Inhibitor drugs | Percentage | 100 | 1 | Heart Failure |
| 25 | Patients underwent echo ejection-fraction procedures | Percentage of patients who received echo ejection-fraction procedures | Percentage | 100 | -1 | Heart Failure |

[K-means clustering](https://www.datanovia.com/learn/machine-learning/clustering/kmeans)

[Partitioning around medoids (PAM)](https://www.datanovia.com/learn/machine-learning/clustering/k-medoids-pam)

[Hierarchical clustering](https://www.datanovia.com/learn/machine-learning/clustering/hierarchical-clustering)
